---
output: github_document
always_allow_html: yes
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%", 
  dpi = 300,
  message = FALSE,
  warning = FALSE
)
```
# mermaidreporting

<!-- badges: start -->
[![CircleCI build status](https://circleci.com/gh/data-mermaid/mermaidreporting.svg?style=svg)](https://circleci.com/gh/data-mermaid/mermaidreporting)
<!-- badges: end -->

The goal of mermaidreporting is to provide utility functions for summarising and visualizing MERMAID data.

## Installation

You can install the development version of mermaidreporting with:

``` r
# install.packages("devtools")
devtools::install_github("data-mermaid/mermaidreporting@development", upgrade = "never")
```

## Data

Data is available from the [`mermaidr`](https://github.com/data-mermaid/mermaidr/tree/package) package. There are details there on authorization and accessing data.

## Usage

All functions in `mermaidreporting` (and `mermaidr`) are of the form `mermaid_*()`, to take advantage of RStudio auto-completion and make them easier to find when you have other packages loaded.

### Data

First, we read in data using `mermaidr`:

```{r}
library(mermaidr)
library(dplyr)

my_projects <- mermaid_get_my_projects() 

xpdc <- my_projects %>%
  filter(name == "XPDC Kei Kecil 2018")

xpdc_fishbelt <- xpdc %>%
  mermaid_get_project_data(method = "fishbelt", data = "sampleevents")
```

## Cleaning data

### Data frame columns

The data from `mermaidr` comes in a form that may seem a little unfriendly to work with! 

For example, in `xpdc_fishbelt`, the mean total biomass data is contained in a *data frame column*:

```{r}
xpdc_fishbelt %>%
  select(biomass_kgha_by_trophic_group_avg)
```

Use `mermaid_clean_columns()` to clean this up and split each trophic group into its own column:

```{r}
library(mermaidreporting)

xpdc_fishbelt %>%
  select(biomass_kgha_by_trophic_group_avg) %>%
  mermaid_clean_columns()
```

This works on Benthic PIT data, as well, and auto-detects any columns in this format to clean!

```{r}
xpdc_benthicpit <- xpdc %>%
  mermaid_get_project_data(method = "benthicpit", data = "sampleevents")

xpdc_benthicpit %>%
  select(percent_cover_by_benthic_category_avg) %>%
  mermaid_clean_columns()
```

### Management rules

Default data from `mermaidr` has specific details about management rules:

```{r}
north_sulawesi_fishbelt <- my_projects %>%
  filter(name == "North Sulawesi (North Minahasa, Sitaro, Tatoareng MPA) Ecological Survey") %>%
  mermaid_get_project_data("fishbelt", "sampleevents")

north_sulawesi_fishbelt %>%
  distinct(management_rules)
```

If you want to look at management rules more broadly, you can clean up the data and group together Partial Restrictions using `mermaid_clean_management_rules`:

```{r}
north_sulawesi_fishbelt %>%
  distinct(management_rules) %>%
  mermaid_clean_management_rules()
```

This groups together the Partial Restrictions rules, tidies up the names, and converts any empty strings to `NA`s for better missing-value handling!

If you want to have a different value, instead of `NA`, you can change it via `missing_value`:

```{r}
north_sulawesi_fishbelt %>%
  distinct(management_rules) %>%
  mermaid_clean_management_rules(missing_value = "Not Specified")
```

## Maps

You can generate map of site locations using `mermaid_map_sites_static()` and `mermaid_map_sites_interactive()`, which produce static and interactive maps, respectively.

```{r map-sites-static, fig.show = "hide"}
library(mermaidreporting)

xpdc_sites <- xpdc %>%
  mermaid_get_project_sites()

mermaid_map_sites_static(xpdc_sites)
```

```{r map-sites-interactive, eval = FALSE}
mermaid_map_sites_interactive(xpdc_sites)
```

Note that this is just a screenshot of the interactive plot, but you can try it out in an interactive R session by running the above code!

### Mapping by a variable

You can also map sites by the value of a variable. For example, to map sites by mean total biomass use `mermaid_map_sites_static()` with `plot_var` set to `biomass_kgha_avg`:

```{r, fig.show = "hide"}
mermaid_map_sites_static(xpdc_fishbelt, biomass_kgha_avg)
```

Or by management rules:

```{r, map-sites-by-management_rules, fig.show = "hide"}
xpdc_fishbelt %>%
  mermaid_clean_management_rules() %>%
  mermaid_map_sites_static(management_rules)
```

## Plotting methods data

### Fish Belt biomass

You can plot Fish Belt biomass, split by a grouping variable (such as trophic group) and compared by another variable (such as reef exposure).

```{r plot-fish-family-biomass-management-rules, fig.show = "hide"}
xpdc_fishbelt %>%
  mermaid_plot_fish_belt_biomass(biomass_kgha_by_trophic_group_avg, reef_exposure)
```

Or by reef zone:

```{r plot-fish-family-biomass-reef-zone, fig.show = "hide"}
xpdc_fishbelt %>%
  mermaid_plot_fish_belt_biomass(biomass_kgha_by_trophic_group_avg, reef_zone)
```

### Benthic PIT percent cover

Similarly, you can plot Benthic PIT percent cover, split by a grouping variable (such as benthic category) and compared by another variable (such as management rules):

```{r, fig.show = "hide"}
xpdc_benthicpit %>%
  mermaid_clean_management_rules(missing_value = "Not Specified") %>%
  mermaid_plot_benthic_pit_percent_cover(percent_cover_by_benthic_category_avg, management_rules)
```

